Ethereum: A Secure Decentralised Generalised Transaction Ledger

### 2014-04-07 Paper.tex
EVM gas cost for memory.

\begin{itemize}
\item {\small SHA3} costs $G_{sha3}$ gas
\item {\small SLOAD} costs $G_{sload}$ gas
\item {\small BALANCE} costs $G_{balance}$ gas
\item {\small SSTORE} costs $d.G_{sstore}$ gas where:
\begin{itemize}
\item $d = 2$ if the new value of the storage is non-zero and the old is zero;
\item $d = 0$ if the new value of the storage is zero and the old is non-zero;
\item $d = 1$ otherwise.
\end{itemize}
\item {\small CALL} costs $G_{call} + g$, where $g$ is the quantity of gas provided; some gas may later be refunded.
\item {\small CREATE} costs $G_{create} + g$ gas, where $g$ is the quantity of gas provided.
\item All other operations cost $G_{step}$ gas.
\end{itemize}

Additionally, when memory is accessed with {\small MSTORE}, {\small MLOAD}, {\small RETURN}, {\small SHA3}, {\small CALLDATA} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it. The memory size must always be a whole number of 32-byte words. Suppose that the highest previously accessed memory word index is $i$, and the new word index is $i'$; an additional $G_{m}(i, i')$ gas must be paid:

\begin{equation}
G_{m}(i, i') \equiv \max(0, \ceil[\Bigg]{\frac{i'+1}{32}} - \ceil[\Bigg]{\frac{i+1}{32}}) G_{memory}
\end{equation}

Originally, with no accesses taken place, the highest accessed index of memory $i$ is defined as $-1$.

@@

\begin{eqnarray}
\Xi(\mathbb{S}, g, I) & \equiv & X(\mathbb{S}, \mathbb{M}, I) \\
\mathbb{M}_g & \equiv & g \\
\mathbb{M}_{pc} & \equiv & 0 \\
\mathbb{M}_M & \equiv & [0, 0, ...] \\
\mathbb{M}_i & \equiv & -1 \\
\mathbb{M}_T & \equiv & []
\end{eqnarray}
\begin{equation}
X(\mathbb{S}, \mathbb{M}, I) \equiv \begin{cases}
(\mathbb{S}, \mathbb{M}, I, []) & \text{if} \quad Z(\mathbb{S}, \mathbb{M}, I)\\
(O(\mathbb{S}, \mathbb{M}, I), \mathbf{o}) & \text{if} \quad \mathbf{o} \neq \varnothing\\
X(O(\mathbb{S}, \mathbb{M}, I)) & \text{otherwise}\\
\end{cases}
\end{equation}

@@

The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, M, i, T)$ which are the gas available, the program counter, the memory contents, the highest used memory index, and the stack contents. The memory contents $\mathbb{M}_M$ are a series of zeroes of size $2^{256}$.

@@

The exceptional halting function $Z$ is defined as:
\begin{equation}
Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \vee \lVert\mathbb{M}_T\rVert < \mathbf{\delta}_w
\end{equation}

@@

The application of rewards to a block involves little more

@@

\begin{equation}
C(\mathbb{S}, \mathbb{M}) \equiv C_M(\mathbb{S}, \mathbb{M}) + \begin{cases}
c_\text{\tiny SSTORE} & \text{if} \quad o = \text{\small SSTORE} \\
G_{call} & \text{if} \quad o = \text{\small CALL}\\
G_{create} & \text{if} \quad o = \text{\small CREATE}\\
G_{sha3} & \text{if} \quad o = \text{\small SHA3}\\
G_{sload} & \text{if} \quad o = \text{\small SLOAD}\\
G_{balance} & \text{if} \quad o = \text{\small BALANCE}\\
G_{stop} & \text{if} \quad o = \text{\small STOP}\\
G_{step} & \text{otherwise}
\end{cases}
\end{equation}

where $c_o$ is specified in the appropriate section below and $C_M$ is the memory cost function, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\mathbb{M}_i$ in order that all accesses reference valid memory whether for read or write; $\mathbb{M}'_i$ is defined as this new maximum number of words of active memory.

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$.} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
&&&& $\mathbb{M}'_R = []$ \\
\midrule
0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] + \mathbb{M}_T[1]$ \\
\midrule
0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] \times \mathbb{M}_T[1]$ \\
\midrule
0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] - \mathbb{M}_T[1]$ \\
\midrule
0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_T[0] \div \mathbb{M}_T[1]\rfloor$ \\
\midrule
0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_T[0] \div \mathbb{M}_T[1]\rfloor$ \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] \mod{ \mathbb{M}_T[1] }$ \\
\midrule
0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] \mod{ \mathbb{M}_T[1] }$ \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] ^ {\mathbb{M}_T[1] }$ \\
\midrule
0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
&&&& $\mathbb{M}'_T[0] \equiv -\mathbb{M}_T[0]$ \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] < \mathbb{M}_T[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] > \mathbb{M}_T[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0c & {\small EQ} & 2 & 1 & Equality comparision. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] = \mathbb{M}_T[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0d & {\small NOT} & 1 & 1 & Simple not operator. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{10s: Bitwise Logic Operations}} \\
\multicolumn{5}{l}{$\mathbb{M}_T[0]_i$ gives the $i$th bit (counting from zero) of $\mathbb{M}_T[0]$} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_T[0]_i \wedge \mathbb{M}_T[1]_i$ \\
\midrule
0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_T[0]_i \vee \mathbb{M}_T[1]_i$ \\
\midrule
0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_T[0]_i \oplus \mathbb{M}_T[1]_i$ \\
\midrule
0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
&&&& $\forall i \in [0..7]: \mathbb{M}'_T[0]_i \equiv \begin{cases} \mathbb{M}_T[1]_{(i + 8\mathbb{M}_T[0])} & \text{if} \quad \mathbb{M}_T[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x20 & {\small SHA3} & 2 & 1 & Compute SHA3-256 hash. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathcal{S}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
\bottomrule
\end{tabular*}

@@

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
\midrule
0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0]$ \\
\midrule
0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[1]$ \\
&&&& $\mathbb{M}'_T[1] \equiv \mathbb{M}_T[0]$ \\
\midrule
0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_M[\mathbb{M}_T[0] \dots (\mathbb{M}_T[0] + 31) ]$ \\
\midrule
0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_T[0] \dots (\mathbb{M}_T[0] + 31) ] \equiv \mathbb{M}_T[1]$ \\
\midrule
0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_T[0] ] \equiv (\mathbb{M}_T[1] \mod 256) $ \\
\midrule
0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}'[I_a]_{storage}[\mathbb{M}_T[0]]$ \\
\midrule
0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
&&&& $\mathbb{S}'[I_a]_{storage}[ \mathbb{M}_T[0] ] \equiv \mathbb{M}_T[1] $ \\
&&&& $c_{\text{\tiny SSTORE}} \equiv \begin{cases}
2G_{sstore} & \text{if} \quad \mathbb{M}_T[1] \neq 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_T[0]] = 0 \\
0 & \text{if} \quad \mathbb{M}_T[1] = 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_T[0]] \neq 0 \\
G_{sstore} & \text{otherwise}
\end{cases}$ \\
\midrule
0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
&&&& $J_{\text{\tiny JUMP}}(\mathbb{M}) \equiv \mathbb{M}_T[0] $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_T[0] & \text{if} \quad \mathbb{M}_T[0] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{pc}$ \\
\midrule
0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{i}$ \\
\midrule
0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{g}$ \\
\bottomrule
\end{tabular*}

>>>>

\begin{itemize}
\item {\small SHA3} costs $G_{sha3}$ gas
\item {\small SLOAD} costs $G_{sload}$ gas
\item {\small BALANCE} costs $G_{balance}$ gas
\item {\small SSTORE} costs $d.G_{sstore}$ gas where:
\begin{itemize}
\item $d = 2$ if the new value of the storage is non-zero and the old is zero;
\item $d = 0$ if the new value of the storage is zero and the old is non-zero;
\item $d = 1$ otherwise.
\end{itemize}
\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of contract body code.
\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of contract initialisation code.
\item All other operations cost $G_{step}$ gas.
\end{itemize}

Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it. See Appendix \ref{app:vm} for a rigourous definition of the EVM gas cost.

@@

\begin{eqnarray}
\Xi(\mathbb{S}, g, I) & \equiv & X(\mathbb{S}, \mathbb{M}, I) \\
\mathbb{M}_g & \equiv & g \\
\mathbb{M}_{pc} & \equiv & 0 \\
\mathbb{M}_M & \equiv & [0, 0, ...] \\
\mathbb{M}_i & \equiv & 0 \\
\mathbb{M}_S & \equiv & []
\end{eqnarray}
\begin{equation}
X(\mathbb{S}, \mathbb{M}, I) \equiv \begin{cases}
(\mathbb{S}, \mathbb{M}, I, []) & \text{if} \quad Z(\mathbb{S}, \mathbb{M}, I)\\
(O(\mathbb{S}, \mathbb{M}, I), \mathbf{o}) & \text{if} \quad \mathbf{o} \neq \varnothing\\
X(O(\mathbb{S}, \mathbb{M}, I)) & \text{otherwise}\\
\end{cases}
\end{equation}

@@

The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, M, i, T)$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents. The memory contents $\mathbb{M}_M$ are a series of zeroes of size $2^{256}$.

@@

The exceptional halting function $Z$ is defined as:
\begin{equation}
Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \vee \lVert\mathbb{M}_S\rVert < \mathbf{\delta}_w
\end{equation}

@@

The application of rewards to a block involves placing in the account of the miner-determined coinbase address additional Ether according to

@@

\begin{equation}
C(\mathbb{S}, \mathbb{M}) \equiv G_{memory}(\mathbb{M}'_i - \mathbb{M}_i) + \begin{cases}
c_\text{\tiny SSTORE} & \text{if} \quad o = \text{\small SSTORE} \\
G_{call} & \text{if} \quad o = \text{\small CALL}\\
G_{create} & \text{if} \quad o = \text{\small CREATE}\\
G_{sha3} & \text{if} \quad o = \text{\small SHA3}\\
G_{sload} & \text{if} \quad o = \text{\small SLOAD}\\
G_{balance} & \text{if} \quad o = \text{\small BALANCE}\\
G_{stop} & \text{if} \quad o = \text{\small STOP}\\
G_{step} & \text{otherwise}
\end{cases}
\end{equation}

where $c_o$ is specified in the appropriate section below and $C_M$ is the memory cost function, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\mathbb{M}_i$ in order that all accesses reference valid memory whether for read or write; $\mathbb{M}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$.} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
&&&& $\mathbb{M}'_R = []$ \\
\midrule
0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] + \mathbb{M}_S[1]$ \\
\midrule
0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] \times \mathbb{M}_S[1]$ \\
\midrule
0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] - \mathbb{M}_S[1]$ \\
\midrule
0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor$ \\
\midrule
0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor$ \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] \mod{ \mathbb{M}_S[1] }$ \\
\midrule
0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] \mod{ \mathbb{M}_S[1] }$ \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] ^ {\mathbb{M}_S[1] }$ \\
\midrule
0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
&&&& $\mathbb{M}'_T[0] \equiv -\mathbb{M}_S[0]$ \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] < \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] > \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0c & {\small EQ} & 2 & 1 & Equality comparision. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0d & {\small NOT} & 1 & 1 & Simple not operator. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{10s: Bitwise Logic Operations}} \\
\multicolumn{5}{l}{$\mathbb{M}_S[0]_i$ gives the $i$th bit (counting from zero) of $\mathbb{M}_S[0]$} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_S[0]_i \wedge \mathbb{M}_S[1]_i$ \\
\midrule
0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_S[0]_i \vee \mathbb{M}_S[1]_i$ \\
\midrule
0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_S[0]_i \oplus \mathbb{M}_S[1]_i$ \\
\midrule
0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
&&&& $\forall i \in [0..7]: \mathbb{M}'_T[0]_i \equiv \begin{cases} \mathbb{M}_S[1]_{(i + 8\mathbb{M}_S[0])} & \text{if} \quad \mathbb{M}_S[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x20 & {\small SHA3} & 2 & 1 & Compute SHA3-256 hash. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathcal{S}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) \div 32 })$ \\
\bottomrule
\end{tabular*}

@@

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
\midrule
0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0]$ \\
\midrule
0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[1]$ \\
&&&& $\mathbb{M}'_T[1] \equiv \mathbb{M}_S[0]$ \\
\midrule
0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_M[\mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 31) \div 32 })$ \\
\midrule
0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ] \equiv \mathbb{M}_S[1]$ \\
&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 31) \div 32 })$ \\
\midrule
0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] ] \equiv (\mathbb{M}_S[1] \mod 256) $ \\
&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \mathbb{M}_S[0] \div 32 })$ \\
\midrule
0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}'[I_a]_{storage}[\mathbb{M}_S[0]]$ \\
\midrule
0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
&&&& $\mathbb{S}'[I_a]_{storage}[ \mathbb{M}_S[0] ] \equiv \mathbb{M}_S[1] $ \\
&&&& $c_{\text{\tiny SSTORE}} \equiv \begin{cases}
2G_{sstore} & \text{if} \quad \mathbb{M}_S[1] \neq 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]] = 0 \\
0 & \text{if} \quad \mathbb{M}_S[1] = 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]] \neq 0 \\
G_{sstore} & \text{otherwise}
\end{cases}$ \\
\midrule
0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
&&&& $J_{\text{\tiny JUMP}}(\mathbb{M}) \equiv \mathbb{M}_S[0] $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_S[0] & \text{if} \quad \mathbb{M}_S[0] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{pc}$ \\
\midrule
0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{i}$ \\
\midrule
0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{g}$ \\
\bottomrule
\end{tabular*}
